We'll start by setting up some package attachments and ontology objects.
These examples use legacy versions of ontologies because certain
terms and relationships were removed in more modern versions, and
our tooling has not kept up with the changes.

```{r setup,echo=FALSE,results="hide",message=FALSE}
library(knitr)
library(ontoProc)
```

```{r setup2, message=FALSE}
library(ontoProc)
library(ontologyPlot)
library(BiocStyle)  # for package references
cl = getOnto("cellOnto", "2021") # for continuity -- 	has_high_plasma_membrane_amount: list
go = getOnto("goOnto", "2021")  # if updated, some assertions will fail...
pr = getOnto("Pronto", "2021")  # important case change
str(cl)
```


## Legacy approaches
  
The following table describes the most up-to-date resources available with `getOnto`.
<!-- use inst/scripts/desc.R to generate all but the 'purpose' which is added by hand -->

```{r lksco}
data(packDesc2023)
kable(packDesc2023[,-c(1,7)])
``` 
    
Other resources are listed in `packDesc202x` and `packDesc2019`.


